---  
title: "scRNA-seq: QC Report"
author: "SysBioLab"
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    fig_width: 12
---

<style type='text/css'>
    body .main-container {
        max-width: 90%;

    }

    .plot_real_size {
      overflow: scroll;
      max-height: 600px;
    }

    .plot_real_size img{
      max-width: none;
      max-height: none;
    }

    embed {
   
  }   
</style>

<style>
.dropdown-content {
  display: none;
}

.dropdown:hover .dropdown-content {
  display: block;
}
</style>

<style>
  .superbigimage{
      overflow-x:scroll;
      white-space: nowrap;
  }

  .superbigimage img{
     max-width: none;
  }


</style>

# `r name` of project `r experiment`

Report of quality of FASTQ files using FastQC and Cell Ranger

## FastQC metrics

```{r setup, echo = FALSE}
library(patchwork)
library(ggplot2)

metric_df <- cellranger_metric_df <- NULL

if(file.size(metrics) > 0) {
  metric_df <- read.table(file = metrics, sep = '\t', header = TRUE)
  if(nrow(metric_df) < 2) {
    metric_df <- NULL
  }
}

if(file.size(cellranger_metrics) > 0) {
  cellranger_metric_df <- read.table(file = cellranger_metrics, sep = '\t', header = TRUE)
  colnames(cellranger_metric_df) <- tolower(gsub("\\.", "_", colnames(cellranger_metric_df)))
  if(nrow(cellranger_metric_df) < 2) {
    cellranger_metric_df <- NULL
  }
}

```

<details>
<summary>**How to interpret**</summary>
<br>

Metrics from the FASTQ files:

- **Number of sequences**: Total number of sequences in the FASTQ file. More sequences means more sequencing depth.
- **Read maximum length**: Maximum read length. It is usual that R2 reads have more length than R1.
- **Read minimum length**: Minimum read legnth. They do not have to be the same as the maximum reading lengths, but they should be consistent with them.
- **GC content**: Guanine-cytosine content. Percentage of nitrogenous bases that are either guanine (G) or cytosine (C). We expect values around 50%.
- **Quality scores - mean quality per base**: Quality is measured with the Phred score (Q), which is typically in the range 2 - 40, but higher and lower values are sometimes used. The higher the Phred score, the higher the quality.
- **Quality scores - mean quality per base in lower quartile**: Same as the previous one, but focusing in the lower quartile.
- **Quality scores - mean quality per base in 10th decile**: Same as the previous one, but focusing in the tenth decile.
- **Quality scores - weigthed quality per sequence**: Same as the previous one, but the mean is calculated by sequence and weighted by sequence length.
- **Mean indeterminations per base**: Mean of the percent of bases at each position with no base call (N).
- **Weigthed read length**: Weighted mean of read length.

</details>

<div class="superbigimage">
```{r initial_total_sequences, results='asis', echo=FALSE, eval = !is.null(metric_df)}

cat("\n\n### **Number of sequences**\n\n")

make_barplot(metric_df, "initial_total_sequences")

```
</div>

<div class="superbigimage">
```{r initial_read_max_length, results='asis', echo=FALSE, eval = !is.null(metric_df)}

cat("\n\n### **Read maximum length**\n\n")

make_barplot(metric_df, "initial_read_max_length")

```
</div>

<div class="superbigimage">
```{r initial_read_min_length, results='asis', echo=FALSE, eval = !is.null(metric_df)}

cat("\n\n### **Read minimum length**\n\n")

make_barplot(metric_df, "initial_read_min_length")

```
</div>

<div class="superbigimage">
```{r initial_gc, results='asis', echo=FALSE, eval = !is.null(metric_df)}

cat("\n\n### **GC content**\n\n")

make_barplot(metric_df, "initial_.gc")

```
</div>

<!-- <div class="superbigimage"> -->
<!-- ```{r initial_mean_indeterminations_per_base, eval=TRUE, results='asis', echo=FALSE, eval = !is.null(metric_df)} -->

<!-- cat("\n\n### **Mean indeterminations per base**\n\n") -->

<!-- make_barplot(metric_df, "initial_mean_indeterminations_per_base") -->

<!-- ``` -->
<!-- </div> -->

<div class="superbigimage">
```{r initial_weigthed_read_length, eval=TRUE, results='asis', echo=FALSE, eval = !is.null(metric_df)}

cat("\n\n### **Weigthed read length**\n\n")

make_barplot(metric_df, "initial_weigthed_read_length")

```
</div>

## Cell Ranger metrics

<details>
<summary>**How to interpret**</summary>
<br>

Metrics from the Cell Ranger alignment/single-cell matrix:

- **Estimated Number of Cells**: Number of barcodes associated with at least one cell.
- **Mean Reads per Cell**: Total number of sequenced reads divided by the number of barcodes associated with cell-containing partitions.
- **Median Genes per Cell**: Median number of genes detected per cell-associated barcode. Detection is defined as the presence of at least 1 UMI count.
- **Valid Barcodes percentage**: Fraction of reads with barcodes that match the whitelist after barcode correction.
- **Sequencing Saturation**: Fraction of reads originating from an already-observed UMI. This is a function of library complexity and sequencing depth. More specifically, this is the fraction of confidently mapped, valid cell-barcode, valid UMI reads that had a non-unique (cell-barcode, UMI, gene).
- **Q30 Bases in Barcode**: Fraction of cell barcode bases with Q-score >= 30, excluding very low quality/no-call (Q <= 2) bases from the denominator.
- **Q30 Bases in RNA Read**: Fraction of RNA read bases with Q-score >= 30, excluding very low quality/no-call (Q <= 2) bases from the denominator. 
- **Q30 Bases in UMI**: Fraction of UMI bases with Q-score >= 30, excluding very low quality/no-call (Q <= 2) bases from the denominator.
- **Reads Mapped to Genome**: Fraction of reads that mapped to the genome.
- **Reads Mapped Confidently to Genome**: Fraction of reads that mapped uniquely to the genome. If a gene mapped to exonic loci from a single gene and also to non-exonic loci, it is considered uniquely mapped to one of the exonic loci.
- **Reads Mapped Confidently to Intergenic Regions**: Fraction of reads that mapped uniquely to an intergenic region of the genome.
- **Reads Mapped Confidently to Intronic Regions**: Fraction of reads that mapped uniquely to an intronic region of the genome.
- **Reads Mapped Confidently to Exonic Regions**: Fraction of reads that mapped uniquely to an exonic region of the genome.
- **Reads Mapped Confidently to Transcriptome**: Fraction of reads that mapped to a unique gene in the transcriptome. The read must be consistent with annotated splice junctions. These reads are considered for UMI counting.
- **Reads Mapped Antisense to Gene**: Fraction of reads confidently mapped to the transcriptome, but on the opposite strand of their annotated gene. A read is counted as antisense if it has any alignments that are consistent with an exon of a transcript but antisense to it, and has no sense alignments. Ideal < 20%. Rates of up to 40% are common for single nuclei samples. Higher fraction of antisense reads may indicate use of an incorrect chemistry type, or an issue with the reference transcriptome.
- **Fraction Reads in Cells**: Fraction of valid-barcode, confidently-mapped reads with cell-associated barcodes.
- **Total Genes Detected**: Number of genes with at least one UMI count in any cell.
- **Median UMI Counts per Cell**: The median number of UMI counts per %s cell-associated barcode.

</details>

<div class="superbigimage">
```{r Estimated_Number_of_Cells, results='asis', echo=FALSE, eval = !is.null(cellranger_metric_df)}

cat("\n\n### **Estimated Number of Cells**\n\n")

make_barplot(cellranger_metric_df, "estimated_number_of_cells")

```
</div>

<div class="superbigimage">
```{r Mean_Reads_per_Cell, results='asis', echo=FALSE, eval = !is.null(cellranger_metric_df)}

cat("\n\n### **Mean Reads per Cell**\n\n")

make_barplot(cellranger_metric_df, "mean_reads_per_cell")

```
</div>

<div class="superbigimage">
```{r Median_Genes_per_Cell, results='asis', echo=FALSE, eval = !is.null(cellranger_metric_df)}

cat("\n\n### **Median Genes per Cell**\n\n")

make_barplot(cellranger_metric_df, "median_genes_per_cell")

```
</div>

<div class="superbigimage">
```{r Valid_Barcodes, results='asis', echo=FALSE, eval = !is.null(cellranger_metric_df)}

cat("\n\n### **Valid Barcodes percentage**\n\n")

make_barplot(cellranger_metric_df, "valid_barcodes")

```
</div>

<div class="superbigimage">
```{r Sequencing_Saturation, results='asis', echo=FALSE, eval = !is.null(cellranger_metric_df)}

cat("\n\n### **Sequencing Saturation**\n\n")

make_barplot(cellranger_metric_df, "sequencing_saturation")

```
</div>

<div class="superbigimage">
```{r Fraction_Reads_in_Cells, results='asis', echo=FALSE, eval = "fraction_reads_in_cells" %in% colnames(cellranger_metric_df)}

cat("\n\n### **Fraction Reads in Cells**\n\n")

make_barplot(cellranger_metric_df, "fraction_reads_in_cells")

```
</div>

<div class="superbigimage">
```{r Total_Genes_Detected, results='asis', echo=FALSE, eval = !is.null(cellranger_metric_df)}

cat("\n\n### **Total Genes Detected**\n\n")

make_barplot(cellranger_metric_df, "total_genes_detected")

```
</div>

<div class="superbigimage">
```{r Median_UMI_Counts_per_Cell, results='asis', echo=FALSE, eval = !is.null(cellranger_metric_df)}

cat("\n\n### **Median UMI Counts per Cell**\n\n")

make_barplot(cellranger_metric_df, "median_umi_counts_per_cell")

```
</div>